---
title: "MFRI by EPA L4 Ecoregion map"
author: "The Nature Conservancy's LANDFIRE team"
date: "10/1/2021"
output: html_document
---


# Goal
Make map of weighted average MFRI per EPA Level 4 Ecoregions with the following schema:

*from Joe Fargione*
Frequent (0-5 years)
Common (6-15 years)
Episodic (16-35 years)
infrequent (36-100 years)
Rare (100+)


# Methods (not the cleanest!)

1. Converted EPA Level 4 Ecoregions Shapefile to Raster, reprojected to LF BpS projection
2. Combined LF200 BpS with ecoregions raster
3. Joined attributes to combined raster
4. Exported attributes to Excel
5. Cleaned attributes in Excel Pivot Table (e.g., retained on the columns I needed, removed NAs as we didn't want them to "contribute" to the average)
6. Moved to R to calculate weighted means using https://datacornering.com/how-to-calculate-weighted-mean-in-r/


# The work

Libraries, read in data
```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)

df<- read_csv("epa_bps_mfri_count.csv")
#View(df)

```

COUNT are the weights, and  MFRI are the values being averaged.  THere should be 967 rows when finished (to match number of ecoregions by L4code)

```{r}
weightedAverages <- df %>%
  group_by(epa_value) %>%
  summarise(mfri = weighted.mean(mfri_all, count))
```

looks OK.  need to create categories.  

Frequent (0-5 years)
Common (6-15 years)
Episodic (16-35 years)
infrequent (36-100 years)
Rare (100+)

```{r}
weightedAverages$labels <- cut(weightedAverages$mfri,
                               breaks = c(-Inf, 5, 15, 35, 100, Inf),
                               labels = c("Frequent (0-5 years)",
                                          "Common (6-15 years)",
                                          "Episodic (16-35 years)",
                                          "Infrequent (36-100 years)",
                                          "Rare (100+)")
)
```




1. read in shapefile
2. join in mfri 
3. map in leaflet!  Try to use different projection than the WGS84!

read in and project shapefile for leaflet 
```{r}
library(sf)
library(sp)
ecoregions <- st_read("epaL4singlePart.shp")
format(object.size(ecoregions), units = "Mb")

ecoregions84 <- st_transform(ecoregions, CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))
format(object.size(ecoregions84), units = "Mb")

outerUS <- st_read("USouter.shp")
outerUS <- st_transform(outerUS, CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))


```

join data

```{r}
ecoregions84mfri <- merge(ecoregions84, weightedAverages, by.x="gridcodesm", by.y = "epa_value")

format(object.size(ecoregions84mfri), units = "Mb")
```

make map

```{r}
library(leaflet)
library(RColorBrewer)
library(viridis)

#make palette
pal <- colorFactor(palette = c('#993404','#d95f0e','#fe9929','#fed98e','#ffffd4'),
                   levels = c("Frequent (0-5 years)",
                              "Common (6-15 years)",
                              "Episodic (16-35 years)",
                              "Infrequent (36-100 years)",
                              "Rare (100+)"))



mfriMap <- leaflet() %>%
  setView(-96, 37.8, 5) %>%
  addProviderTiles(providers$Esri.WorldShadedRelief) %>%
  addPolygons(data = ecoregions84mfri,
              fillColor = ~pal(labels),
              fillOpacity = 0.9,
              weight = 0.5,
              color = "#a1a3a6",
              opacity = 0.8) %>%
  addLegend(pal = pal,
            values = ecoregions84mfri$labels,
            position = "bottomleft",
            title = "Historical Mean Fire Return Interval",
            opacity = 1)

mfriMap

library(htmlwidgets)
saveWidget(mfriMap, 'mfriMap.html')

```







